A Gaussian process model of human e\-lect\-ro\-cort\-i\-co\-graph\-ic data

### 2020-04-15 paper/main.tex
fixed typo
proposed modifications to address reviewer concerns

\begin{abstract}
  We present a model-based method for inferring full-brain neural
  activity at millimeter-scale spatial resolutions and
  millisecond-scale temporal resolutions using standard human
  intracranial recordings.  Our approach assumes that different
  people's brains exhibit similar correlational structure, and that
  activity and correlation patterns vary smoothly over space.  One can
  then ask, for an arbitrary individual's brain: given recordings from
  a limited set of locations in that individual's brain, along with
  the observed spatial correlations learned from other people's
  recordings, how much can be inferred about ongoing activity at
  \textit{other} locations throughout that
  individual's brain?  We show that our approach generalizes across
  people and tasks, thereby providing a person- and task-general means
  of inferring high spatiotemporal resolution full-brain neural
  dynamics from standard low-density intracranial recordings.
  \\\\
  \footnotesize{\textbf{Keywords: Electrocorticography (ECoG),
      intracranial electroencephalography (iEEG), local field
      potential (LFP), epilepsy, maximum likelihood estimation,
      Gaussian process regression}}
\end{abstract}

@@

Although both datasets we examined provide good full-brain coverage (when
considering data from every patient), electrodes were not sampled uniformly
throughout the brain.  For example, in our patient population, electrodes are
more likely to be implanted in regions like the medial temporal lobe (MTL), and
are rarely implanted in occipital cortex (Fig.~\ref{fig:density}A, B).
Separately for each dataset, for each voxel in the 4~mm$^3$ voxel MNI152 brain,
we computed the proportion of electrodes in the dataset that were contained
within a 20 MNI unit radius sphere centered on that voxel.  We defined the
\textit{density} at that location as this proportion. Across Datasets 1 and 2,
the electrode placement densities were similar (correlation by voxel: $r = 0.6,
p < 10^{-10}$).  We wondered whether regions with good covererage might be
associated with better reconstruction accuracy. For example,
Figures~\ref{fig:corrmap}C and D indicate that some electrodes in the MTL (which
tends to be relatively densely sampled) have relatively high reconstruction
accuracy, and occipital electrodes (which tends to be relatively sparsely
sampled) tend to have relatively low reconstruction accuracy.  To test whether
this held more generally across the entire brain, for each dataset we computed
the electrode placement density for each electrode from each patient (using the
proportion of \textit{other} patients' electrodes within 20 MNI units of the
given electrode).  We then correlated these density values with the
across-patient reconstruction accuracies for each electrode.  We found no
reliable correlation between reconstruction accuracy and density for Dataset 1
($r = 0.05, p = 0.70$) and a reliable negative correlation for Dataset 2 ($r =
-0.21, p = 0.05$).  This suggests that the reconstruction accuracies we
observed are \textit{not} driven solely by sampling density, but rather may also
reflect higher order properties of neural dynamics such as functional
correlations between distant voxels~\citep{BetzEtal17b}.

Prior work in humans and animals has shown that the spatial profile of the local
field potential differs by frequency band~\citep[e.g., with respect to volume
conductance properties and contribution to the local field
potential;][]{BuzsEtal12, FrieEtal07, CronEtal11}.  For example, lower frequency
components of the local field potential tend to have higher power and extend
further in space than high-frequency components~\citep[e.g., ][]{MillEtal07,
MannEtal09}.  We wondered whether the reconstructions we observed might be
differently weighting or considering the contributions of activity at different
frequency bands.  We therefore examined a range of frequency bands ($\delta$:
2--4 Hz; $\theta$: 4--8 Hz; $\alpha$: 8--12 Hz; $\beta$: 12--30 Hz; $\gamma_L$:
30--60 Hz; and $\gamma_H$: 60--100 Hz), along with a measure of broadband (BB)
power. We used second-order Butterworth bandpass filters to compute the activity
patterns within each narrow frequency band.  We defined broadband power as the
mean height of a linear robust regression fit in log-log space to the order 4
Morelet wavelet-computed power spectrum at 50 log-spaced frequencies from from
2--100 Hz~\citep{MannEtal09}. We then repeated our within-subject and
across-subject cross-validated reconstruction accuracy tests (analogous to
Fig.~\ref{fig:corrmap}) separately for each frequency band
(Fig.~\ref{fig:frequency}). (We also carried out a similar analysis on the
Hilbert transform-computed spectral power within each narrow band; see
Fig.~\freqpower.) Across both datasets, we found that our approach is best at
reconstructing patterns of broadband activity (right-most bars in
Figs.~\ref{fig:frequency}A and D), a correlate of population firing
rate~\citep{MannEtal09}.  We also achieved good reconstruction accuracy within
each narrow frequency band (Figs.~\ref{fig:frequency} and \freqpower).  Activity
at lower frequencies ($\delta$, $\theta$, $\alpha$, and $\beta$) tended to be
reconstructed better than high-frequency patterns ($\gamma_L$ and $\gamma_H$),
with reconstruction accuracy peaking in the $\theta$ band.  Overall, these
results indicate that our approach is able to accurately recover information
within the 2--100 Hz range.

@@

Taken together, the frequency-specific information maps suggest a potential new
interpretation of many of the above previously reported findings.  Prior work
has largely treated region-specific narrowband and broadband activity as an
indictor that activity at those frequency ranges reflects that the given region
is representing or supporting a particular function.  Our work suggests an
alternative interpretation that when we observe a particular neural pattern in a
particular brain region, it may instead reflect how that region is is
transmitting information to the rest of the brain via signalling at the given
frequency range.

\section*{Discussion}
Are our brain's networks static or dynamic?  And to what
extent are the network properties of our brains stable across people and tasks?
One body of work suggests that our brain's \textit{functional} networks are
dynamic~\citep[e.g., ][]{MannEtal18, OwenEtal19}, person-specific~\citep[e.g.,
][]{FinnEtal15}, and task-specific~\citep[e.g., ][]{Turk13}.  In contrast,
although the gross anatomical structure of our brains changes meaningfully over
the course of years as our brains develop, on the timescales of typical
neuroimaging experiments (i.e., hours to days) our anatomical networks are
largely stable~\citep[e.g., ][]{CaseEtal00}.  Further, many aspects of brain
anatomy, including white matter structure, are largely preserved across
people~\citep[e.g., ][]{TalaTour88, JahaEtal13, MoriEtal08}. There are several
possible means of reconciling this apparent inconsistency between dynamic
person- and task-specific functional networks versus stable anatomical networks.
For example, relatively small magnitude anatomical differences across people may
be reflected in reliable functional connectivity differences.  Along these
lines, one recent study found that diffusion tensor imaging (DTI) structural
data is similar across people, but may be used to predict person-specific
resting state functional connectivity data~\citep{BeckEtal18}.  Similarly, other
work indicates that task-specific functional connectivity may be predicted by
resting state functional connectivity data~\citep{ColeEtal16, TavoEtal16}.
Another (potentially complementary) possibility is that our functional networks
are constrained by anatomy, but nevertheless exhibit (potentially rapid)
task-dependent changes~\citep[e.g., ][]{SporBetz16}.

@@

@@

@@

By providing a means of reconstructing full-brain activity patterns, the
SuperEEG approach maps ECoG recordings from different patients into a common
neural space, despite that different patients' electrodes were implanted in
different locations. This feature of our approach enables across-patient ECoG
studies, analogous to across-patient fMRI studies~\citep[e.g.,][]{HaxbEtal01,
NormEtal06, HaxbEtal11}. Whereas the focus of this manuscript is to specifically
evaluate which aspects of neural activity patterns SuperEEG recovers well (or
poorly), in parallel work we are training across-patient classifiers by
leveraging the common neural spaces obtained by applying SuperEEG to
multi-patient ECoG data.  For example, we have shown that SuperEEG-derived
activity patterns may be used to accurately predict psychiatric conditions such
as depression~\citep{ScanEtal20}.  Analogous approaches could in principle be
used to develop improved brain-computer interfaces and/or to carry out other
analyses that would benefit from high spatiotemporal resolution full-brain data
in individuals, projected into a common ECoG space across people.

>>>>

\begin{abstract}
  We present a model-based method for inferring full-brain neural
  activity at millimeter-scale spatial resolutions and
  millisecond-scale temporal resolutions using standard human
  intracranial recordings.  Our approach makes the simplifying
  assumptions that different people's brains exhibit similar
  correlational structure, and that activity and correlation patterns
  vary smoothly over space.  One can then ask, for an arbitrary
  individual's brain: given recordings from a limited set of locations
  in that individual's brain, along with the observed spatial
  correlations learned from other people's recordings, how much can be
  inferred about ongoing activity at \textit{other} locations
  throughout that individual's brain?  We show that our approach
  generalizes across people and tasks, thereby providing a person- and
  task-general means of inferring high spatiotemporal resolution
  full-brain neural dynamics from standard low-density intracranial
  recordings.
  \\\\
  \footnotesize{\textbf{Keywords: Electrocorticography (ECoG),
      intracranial electroencephalography (iEEG), local field
      potential (LFP), epilepsy, maximum likelihood estimation,
      Gaussian process regression}}
\end{abstract}

@@

Although both datasets we examined provide good full-brain coverage (when
considering data from every patient), electrodes were not sampled uniformly
throughout the brain.  For example, in our patient population, electrodes are
more likely to be implanted in regions like the medial temporal lobe (MTL), and
are rarely implanted in occipital cortex (Fig.~\ref{fig:density}A, B).
Separately for each dataset, for each voxel in the 4~mm$^3$ voxel MNI152 brain,
we computed the proportion of electrodes in the dataset that were contained
within a 20 MNI unit radius sphere centered on that voxel.  We defined the
\textit{density} at that location as this proportion. Across Datasets 1 and 2,
the electrode placement densities were similar (correlation by voxel: $r = 0.6,
p < 10^{-10}$).  We wondered whether regions with good coverage might be
associated with better reconstruction accuracy. For example,
Figures~\ref{fig:corrmap}C and D indicate that some electrodes in the MTL (which
tends to be relatively densely sampled) have relatively high reconstruction
accuracy, and occipital electrodes (which tends to be relatively sparsely
sampled) tend to have relatively low reconstruction accuracy.  To test whether
this held more generally across the entire brain, for each dataset we computed
the electrode placement density for each electrode from each patient (using the
proportion of \textit{other} patients' electrodes within 20 MNI units of the
given electrode).  We then correlated these density values with the
across-patient reconstruction accuracies for each electrode.  We found no
reliable correlation between reconstruction accuracy and density for Dataset 1
($r = 0.05, p = 0.70$) and a reliable negative correlation for Dataset 2 ($r =
-0.21, p = 0.05$).  This suggests that the reconstruction accuracies we
observed are \textit{not} driven solely by sampling density, but rather may also
reflect higher order properties of neural dynamics such as functional
correlations between distant voxels~\citep{BetzEtal17b}.

Prior work in humans and animals has shown that the spatial profile of the local
field potential differs by frequency band~\citep[e.g., with respect to volume
conductance properties and contribution to the local field
potential;][]{BuzsEtal12, FrieEtal07, CronEtal11}.  For example, lower frequency
components of the local field potential tend to have higher power and extend
further in space than high-frequency components~\citep[e.g., ][]{MillEtal07a,
MannEtal09}.  We wondered whether the reconstructions we observed might be
differently weighting or considering the contributions of activity at different
frequency bands.  We therefore examined a range of frequency bands ($\delta$:
2--4 Hz; $\theta$: 4--8 Hz; $\alpha$: 8--12 Hz; $\beta$: 12--30 Hz; $\gamma_L$:
30--60 Hz; and $\gamma_H$: 60--100 Hz), along with a measure of broadband (BB)
power. We used second-order Butterworth bandpass filters to compute the activity
patterns within each narrow frequency band.  We defined broadband power as the
mean height of a linear robust regression fit in log-log space to the order 4
Morelet wavelet-computed power spectrum at 50 log-spaced frequencies from from
2--100 Hz~\citep{MannEtal09}. We then repeated our within-subject and
across-subject cross-validated reconstruction accuracy tests (analogous to
Fig.~\ref{fig:corrmap}) separately for each frequency band
(Fig.~\ref{fig:frequency}). (We also carried out a similar analysis on the
Hilbert transform-computed spectral power within each narrow band; see
Fig.~\freqpower.) Across both datasets, we found that our approach is best at
reconstructing patterns of broadband activity (right-most bars in
Figs.~\ref{fig:frequency}A and D), a correlate of population firing
rate~\citep{MannEtal09}.  We also achieved good reconstruction accuracy within
each narrow frequency band (Figs.~\ref{fig:frequency} and \freqpower).  Activity
at lower frequencies ($\delta$, $\theta$, $\alpha$, and $\beta$) tended to be
reconstructed better than high-frequency patterns ($\gamma_L$ and $\gamma_H$),
with reconstruction accuracy peaking in the $\theta$ band.  Overall, these
results indicate that our approach is able to accurately recover information
within the 2--100 Hz range.

@@

\section*{Discussion} Are our brain's networks static or dynamic?  And to what
extent are the network properties of our brains stable across people and tasks?
One body of work suggests that our brain's \textit{functional} networks are
dynamic~\citep[e.g., ][]{MannEtal18, OwenEtal19}, person-specific~\citep[e.g.,
][]{FinnEtal15}, and task-specific~\citep[e.g., ][]{Turk13}.  In contrast,
although the gross anatomical structure of our brains changes meaningfully over
the course of years as our brains develop, on the timescales of typical
neuroimaging experiments (i.e., hours to days) our anatomical networks are
largely stable~\citep[e.g., ][]{CaseEtal00}.  Further, many aspects of brain
anatomy, including white matter structure, are largely preserved across
people~\citep[e.g., ][]{TalaTour88, JahaEtal13, MoriEtal08}. There are several
possible means of reconciling this apparent inconsistency between dynamic
person- and task-specific functional networks versus stable anatomical networks.
For example, relatively small magnitude anatomical differences across people may
be reflected in reliable functional connectivity differences.  Along these
lines, one recent study found that diffusion tensor imaging (DTI) structural
data is similar across people, but may be used to predict person-specific
resting state functional connectivity data~\citep{BeckEtal18}.  Similarly, other
work indicates that task-specific functional activations may be predicted by
resting state functional connectivity data~\citep{ColeEtal16, TavoEtal16}.
Another (potentially complementary) possibility is that our functional networks
are constrained by anatomy, but nevertheless exhibit (potentially rapid)
task-dependent changes~\citep[e.g., ][]{SporBetz16}.  This prior work differs
from ours in a number of ways.  For example, fMRI data has substantially higher
spatial resolution than (raw) ECoG data, and fMRI data have nearly complete
spatial overlap across participants whereas ECoG data have minimal spatial
overlap across participants.  Nevertheless, our work draws inspiration from
those studies in that we also attempt to estimate held-out activity patterns
across people and tasks.

@@

Another important finding is that SuperEEG-based reconstructions accurately
recover activity patterns at a broad range of frequencies (as well as broadband
patterns).  However, brain networks differed in how informative they were about
activity within each frequency band.  Prior work has largely treated
region-specific narrowband and broadband activity as an indicator that activity
at those frequency ranges reflects that the given region is representing or
supporting a particular function.  Our work suggests a complementary
interpretation that when we observe a particular neural pattern in a particular
brain region, it may instead (or in addition) reflect how that region is
transmitting information to the rest of the brain via signalling at the given
frequency range.

@@

A second potential limitation of our approach is that it relies on
ECoG data from epilepsy patients.  Recent work comparing functional
correlations in epilepsy patients (measured using ECoG) and healthy
individuals (measured using fMRI) suggests that there are gross
similarities between these populations~\citep[e.g.,][]{ReddEtal18,
  KucyEtal18}.  Nevertheless, because all of the patients we examined
have drug-resistant epilepsy, it remains uncertain how generally the
findings reported here might apply more broadly to the population at
large (e.g., non-clinical populations).

@@

By providing a means of reconstructing full-brain activity patterns, the
SuperEEG approach maps ECoG recordings from different patients into a common
neural space, despite that different patients' electrodes were implanted in
different locations. This feature of our approach enables across-patient ECoG
studies, analogous to across-subject fMRI studies~\citep[e.g.,][]{HaxbEtal01,
NormEtal06, HaxbEtal11}. Whereas the focus of this manuscript is to specifically
evaluate which aspects of neural activity patterns SuperEEG recovers well (or
poorly), in parallel work we are training across-patient classifiers by
leveraging the common neural spaces obtained by applying SuperEEG to
multi-patient ECoG data.  For example, we have shown that SuperEEG-derived
activity patterns may be used to accurately predict psychiatric conditions such
as depression~\citep{ScanEtal20}.  Analogous approaches could in principle be
used to develop improved brain-computer interfaces and/or to carry out other
analyses that would benefit from high spatiotemporal resolution full-brain data
in individuals, projected into a common ECoG space across people.
